---
title: "Surfaces"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: true
---

<br>

In this section we continue to ask questions about the distribution of Plasmodesmata (or similar types of annotations along a given model). We continue to use the output of the Spatial control plugin shown in the [Distributions section](#distributions.html). Since we previously detected a bias in the distribution of Plasmodesmata, strongly hinting at the presence of spatial clustering, we then quantified the number of these clusters of plasmodesmata we now ask: **What proportions of the cellular surface do these clusters occupy**

<br>

# Exploring the data

```{r, include = FALSE}

# THIS SECTION IS HIDDEN AND ONLY REQUIRED BECAUSE KNITTING TREATS R MARKDOWNS INDEPENDENTLY so it is not possible to call variables present in another markdown

library(tidyverse)
library(broom)

setwd('./Data_individual_cells')
clusters <- read.csv("clusters.csv")

```

<br>

```{r, message = FALSE, warning= FALSE}

# we loook at one one cell as an example

cluster_example <- clusters %>% filter(DatasetFilename == "170314_Col_HD_R20_339-381um_DNN" & Cell == "PPP1-EN")

# we are going to largely replicate the workings of the fviz_clust function used in clusters section. There these steps were done implicitly by the functions but here we make the steps apparent and we extract data

# we first perform a pca, reducing our data to a 2D situation
# we scale the data to match what was done by the fviz_clust function. Note that this is not strictly necessary but more for graphical matching purposes. Clusters can be recognised in both cases. The scaling can also be turned off in the fviz_clust command. Either way the scaling effect seems negligible

cluster_pca <- cluster_example %>% select(X_units, Y_units, Z_units) %>% prcomp(scale. = TRUE)

# we can check that the PCA is not causing a signicant loss of information

tidy(cluster_pca, matrix = "pcs") %>% print() 
# more than 90% retained in all cases in the first two dimentions

# we attach the pca columns contained in pca object to the original data
# 1 and 2 only take the first two PCAs, we are not attaching third PCA

cluster_example <- cbind(cluster_example, cluster_pca$x[, 1:2]) %>% 
  # we also order the numbers of clusters for that well so that they will appear in a logical order in the legend 
  mutate(.cluster = factor(.cluster, levels = c("1", "2", "3", "4", "5", "6","7","8", "9", "10", "11"))) %>% 
  mutate(.class = factor(.class, levels = c("1", "2", "3", "4", "5", "6","7","8", "9", "10", "11", "12", "13", "14")))

# we use now use the chull function
# it finds the rows that contain the points (and the coordinates) that describe the edges of the poligon that could be drawn around each cluster
# to isolate those rows we use the function slice
# we group by cluster to obtain the single areas or we don't to obtain the full surface of the cell 

cluster_chull_sil <- cluster_example %>% 
  group_by(.cluster) %>% 
  slice(chull(PC1, PC2))

cluster_chull_class <- cluster_example %>% 
  group_by(.class) %>% 
  slice(chull(PC1, PC2))

total_chull <- cluster_example %>% 
  slice(chull(PC1, PC2))

# we visualise the results, in this case for the silhouette method

ggplot(cluster_example, aes(PC1, PC2)) + theme_bw() +
  geom_point(data = cluster_example, aes(colour=.cluster, fill=.cluster)) +
  geom_polygon(data = total_chull, alpha = 0.09) +
  geom_polygon(data = cluster_chull_sil, aes(colour=.cluster, fill = .cluster), alpha = 0.3) +
  xlab("PC1 (78.5%)") + ylab("PC2 (20.5%)") + 
  theme(axis.title.x = element_text(size = 15), 
        axis.title.y = element_text(size = 15), 
        axis.text.x = element_text(size = 15), 
        axis.text.y = element_text(size = 15),
        strip.text.x = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15)) +
  labs(fill = "Silhouette \n clusters", colour = "Silhouette \n clusters")

# you will notice that this is nearly identical to the fviz function results
# by using .class instead you would get the results according to the McLust function

# now we want to calculate areas
# we  created a small function that uses the areapl from the splancs library

cha <- function(x,y){
  i <- chull(x,y)
  return(splancs::areapl(cbind(x[i],y[i])))
}

cluster_example %>% 
  mutate(total_area = cha(PC1, PC2)) %>% 
  group_by(.cluster) %>% 
  summarise(area = cha(PC1, PC2), total_area = unique(total_area)) %>%
  head()

# do note that the units of these values are not strictyly speaking conventional surface units as the coordinates underwent a pca. We address this aspect in following steps

```

<br>

# Compiling the data

<br>

```{r,  message = FALSE, warning= FALSE}

# to process multiple cells in one go we write a larger function encompassing the steps described above 

calculate_areas_sil <- function(x){
  
  # Run PCA
  x_pca <- x %>% 
    select(X_units, Y_units, Z_units) %>% 
    prcomp(scale. = TRUE)
  
  # bind first two PCs
  x <- cbind(x, x_pca$x[, 1:2])
  
  # calculate area of each cluster (here we do it for the silhouette method)
  area_of_clusters <- x %>% 
    group_by(.cluster) %>% 
    summarise(area = cha(PC1, PC2))
  
  # add total area
  area_of_clusters$total_area <- cha(x$PC1, x$PC2)
  
  #add sum of the areas of clusters
  area_of_clusters$sum_area <- sum(area_of_clusters$area)
  
  #percentage conductive surface (of the total)
  area_of_clusters$perc_tot <- area_of_clusters$sum_area/area_of_clusters$total_area*100
  
  #percentage of surface of each cluster 
  area_of_clusters$perc <- area_of_clusters$area/area_of_clusters$total_area*100
  
  return(area_of_clusters)
}


# a nearly identical second function is created for the Mclust method

calculate_areas_class <- function(x){
  
  x_pca <- x %>% 
    select(X_units, Y_units, Z_units) %>% 
    prcomp(scale. = TRUE)
  x <- cbind(x, x_pca$x[, 1:2])
  
  # we use .class instead
  area_of_clusters <- x %>% 
    group_by(.class) %>% 
    summarise(area = cha(PC1, PC2))
  
  area_of_clusters$total_area <- cha(x$PC1, x$PC2)
  
  area_of_clusters$sum_area <- sum(area_of_clusters$area)
  
  area_of_clusters$perc_tot <- area_of_clusters$sum_area/area_of_clusters$total_area*100
  
  area_of_clusters$perc <- area_of_clusters$area/area_of_clusters$total_area*100
  
  return(area_of_clusters)
}

# we apply it and we nest the results

clusters_processed <- clusters %>% 
  group_by(DatasetFilename, Genotype, Interface, Cell) %>% 
  nest() %>% 
  mutate(cluster_sil_areas = lapply(data, calculate_areas_sil)) %>% 
  mutate(cluster_class_areas = lapply(data, calculate_areas_class))

# we then unnest separately the results for either method

area_for_method_sil <- clusters_processed %>% 
  unnest(cluster_sil_areas) %>% 
  mutate(Method = "Silhouette")

area_for_method_class <- clusters_processed %>% 
  unnest(cluster_class_areas) %>% 
  mutate(Method = "Mclust")

# we then merge them

clusters_unnested <- bind_rows(area_for_method_class, area_for_method_sil)

```
<br>

# Calculating the median total conductive surface of an interface (as %)

<br>

```{r}

# we now ask how much of the total surface (the gray shaded area we calculated in previous steps) is occupied by PD clusters

# for plotting you need to isolate the unique value for each cell (otherwise each cluster will be counted as a point altering the shape of distribution).
#the easiest way is to subset the dataset only for relevant columns and then use unique

ggplot(data=unique(clusters_unnested[, c("Genotype", "Interface", "DatasetFilename", "Cell", "perc_tot", "Method")]), aes(x= Method, y=perc_tot, colour= Genotype, fill= Genotype)) + theme_bw() +
  geom_violin(alpha=0.5)+ 
  ggbeeswarm::geom_quasirandom(aes(group=Genotype), colour="black", fill="black", size= 2, alpha=0.5, width=0.1, show.legend = FALSE, dodge.width = 0.9) +
  stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,
               geom = "crossbar", size = 0.5, width = 0.3, alpha=1)  + 
  labs(y = "% of interface surface overall occupied by PD clusters") +
  theme(axis.title.x = element_text(size = 15), 
        axis.title.y = element_text(size = 15), 
        axis.text.x = element_text(size = 15), 
        axis.text.y = element_text(size = 15),
        strip.text.x = element_text(size = 15), 
        legend.title = element_text(size = 15),
        legend.text = element_text(size = 15)) +
  scale_color_manual(values=c("#8B8B83")) +
  scale_fill_manual(values=c("#8B8B83")) +
  facet_grid(~Interface) 

```

<br>

# Calculating the median surface of individual clusters (as %)

<br>

```{r,  message = FALSE, warning= FALSE}

# to calculate the surface of individual clusters independently we don't need to subset the data

ggplot(clusters_unnested, aes(x= Method, y=perc, colour= Genotype, fill= Genotype)) + theme_bw() +
  geom_violin(alpha=0.5)+ 
  ggbeeswarm::geom_quasirandom(aes(group=Genotype), colour="black", fill="black", size= 2, alpha=0.5, width=0.1, show.legend = FALSE, dodge.width = 0.9) +
  stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,
               geom = "crossbar", size = 0.5, width = 0.3, alpha=1)  + 
  labs(y = "% of interface surface occupied by individual PD clusters") +
  theme(axis.title.x = element_text(size = 15), 
        axis.title.y = element_text(size = 15), 
        axis.text.x = element_text(size = 15), 
        axis.text.y = element_text(size = 15),
        strip.text.x = element_text(size = 15), 
        legend.title = element_text(size = 15),
        legend.text = element_text(size = 15)) +
  scale_color_manual(values=c("#8B8B83")) +
  scale_fill_manual(values=c("#8B8B83")) +
  facet_grid(~Interface) +
  scale_y_log10()

```

<br>

# Calculating the actual surface of the interface

<br>
